---
title: "DHGLMs"
author: "Tali Caspi"
date: "2025-07-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(tidyverse); library(brms); library(DHARMa.helpers); library(coda); library(ggpubr)
library(bayestestR); library(purrr); library(broom.mixed)

# set colors
colors <- c("#AF8566", "#B5C861", "#79ACBD", "#E4DECE","#ECBD95", "#0E075A")

# Make custom theme
theme_custom <- function() {
  theme_classic()+
  theme(panel.grid.minor = element_blank(),
        strip.text = element_text(size = 12, face = "bold"),
        plot.background = element_rect(fill = "white", color = NA),
        plot.title = element_text(size = 7),
        strip.background = element_blank(),
        axis.text.y = element_text(size=10, color="black"),
        axis.text.x = element_text(size=10, color="black"),
        axis.title.y = element_text(size=12, color="black"),
        axis.title.x = element_text(size=12, color="black"))}
```

# Load and Format data

```{r}
data <- read.csv("Data/Q1_Summary_Firsts_DummyRows_Max_07282025.csv") %>%  # first visits only
  filter(Species != "Gray Fox" & Species != "Red Fox") %>%   # remove foxes
  filter(Location != "Lake Chabot Dumpster") %>%  # remove Lake Chabot
  filter(Dummy != "dummy") # remove trials with no recorded data

data.RT <- data %>% 
  mutate(OffCamera = ifelse(WasSolveOffCamera == "Yes", "Yes", "No")) %>% # make column if animals was ever off camera
  mutate(PropVigilantMax_RT = ifelse(PropVigilantMax_RT == 1, 0.999, PropVigilantMax_RT)) %>%  # change prop vigilant 1's to 0.999
  mutate(LookAtCamera_RT = ifelse(LookAtCamera_RT == "Yes", 1, 0)) %>% # code Yes as 1 and No as 0
  mutate(Touch_RT = ifelse(Touch_RT == "Yes", 1, 0)) %>%  # code Yes as 1 and No as 0
  
  # separate raccoons into two groups
  mutate(Species = case_when(
    Species == "Raccoon" & Condition == "Alone" ~ "Raccoon.Alone",
    Species == "Raccoon" & Condition == "Group" ~ "Raccoon.Group",
    TRUE ~ Species))
  
data.PS <- data.RT %>% 
  filter(Solve_PS != "DidNotAttempt")  # remove no attempts at solve 
```

# Summary Stats

## Risk-Taking

```{r}
nrow(data.RT) # all observations
table(data.RT$Species)

RT.species.means <- data.RT %>% 
  group_by(Species) %>% 
  summarize(meanPropVig = mean(PropVigilantMax_RT, na.rm=T),
            meanTotExp = mean(TotalExplorationMax_RT, na.rm=T),
            total.look = sum(LookAtCamera_RT, na.rm=T),
            total.touch = sum(Touch_RT, na.rm=T),
            total.trials = n()) %>% 
  mutate(p.look = total.look/total.trials,
         p.touch = total.touch/total.trials)

RT.species.means
```

## Problem-Solving

```{r}
nrow(data.PS) # all problem-solving attempts
table(data.PS$Species) # observations per species

PS.species.means.all <- data.PS %>% 
  group_by(Species) %>% 
  summarize(meanExpDiv = mean(ExploratoryDiversityMax_PS, na.rm=T),
            meanWorkTime = mean(WorkTimeMax_PS, na.rm=T),
            meanActRate = mean(ActivityRateMax_PS, na.rm=T),
            total.solves = sum(Solve_PS_Binary,na.rm=T),
            total.attempts = n()) %>% 
  mutate(p.Solve = total.solves/total.attempts)

PS.species.means
```

# Q1) How do species differ in their risk-taking and problem-solving behavior?

## Set Priors

```{r}
base_priors <- c(
  prior(normal(0, 1), class = "b"),   # fixed effects        
  prior(normal(0, 1.5), class = "Intercept"),  # intercept
  prior(exponential(1), class = "sd")) # random effect

base_priors_zi <-  c(
  prior(normal(0,1), class = "b"),                    
  prior(normal(0,1.5), class = "Intercept"),            
  prior(normal(0,1.5), class = "Intercept", dpar = "zi"), # set prior for zi
  prior(exponential(1), class = "sd"))

base_priors_ln <- c(
  prior(normal(0, 1), class = "b"),      
  prior(normal(0, 2), class = "Intercept"), 
  prior(exponential(1), class = "sd"))
```

## Construct Models

```{r}
set.seed(123)

## RISK TAKING MODELS

# Proportion of time spent vigilant
mod.prop.vig <- brm(bf(PropVigilantMax_RT ~ OffCamera + (1 | Species),
                       zi ~ (1 | Species)),
  family = zero_inflated_beta(), data = data.RT, # zero-inflated beta regression
  prior = base_priors_zi,
  cores = 4, chains = 4,  iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.99, max_treedepth=15))

# Look at camera
mod.look.cam <- brm(LookAtCamera_RT ~ 1 + (1 | Species),
                    family = bernoulli(link = "logit"), data = data.RT, # bernoulli
                    chains = 4,  cores = 4,  iter = 4000,
                    prior = c(prior(normal(0, 1.5), class = "Intercept"),
                              prior(exponential(1), class = "sd")),
                    control = list(adapt_delta = 0.99, max_treedepth=15))

# Touch puzzle
mod.touch <- brm(Touch_RT ~  (1 | Species),
                    family = bernoulli(link = "logit"), data = data.RT, # bernoulli
                    chains = 4,  cores = 4,  iter = 4000,
                    prior = c(prior(normal(0, 1.5), class = "Intercept"),
                              prior(exponential(1), class = "sd")),
                 control = list(adapt_delta = 0.99, max_treedepth=15))

# Total exploration (number of behaviors exhibited in interactions with puzzle)
data.RT_outlier <- data.RT %>% 
  filter(TotalExplorationMax_RT < 113) # remove one outlier with very high value for total exploration

mean_val <- mean(data.RT_outlier$TotalExplorationMax_RT)
var_val <- var(data.RT_outlier$TotalExplorationMax_RT)
cat("Mean:", mean_val, " | Variance:", var_val) # over-dispersed so cannot use poisson distribution -> use negative binomial

mod.tot.exp <- brm(TotalExplorationMax_RT ~ OffCamera + (1 | Species),
                   family = negbinomial(),  data = data.RT_outlier, # negative binomial
                   chains = 4,  cores = 4,  iter = 4000,
                   prior = base_priors_ln,
                   control = list(adapt_delta = 0.99, max_treedepth=15))


## PROBLEM-SOLVING MODELS

# Exploratory diversity
hist(data.PS$ExploratoryDiversityMax_PS)
mean_val <- mean(data.PS$ExploratoryDiversityMax_PS)
var_val <- var(data.PS$ExploratoryDiversityMax_PS)
cat("Mean:", mean_val, " | Variance:", var_val) # look at dispersion

mod.exp.div <- brm(ExploratoryDiversityMax_PS | trunc(lb = 1) ~ OffCamera + (1 | Species),
                   family = poisson(),  data = data.PS, # poisson
                   chains = 4,  cores = 4,  iter = 4000,
                   prior = base_priors,
                   control = list(adapt_delta = 0.99, max_treedepth=15))

# Work time
hist(log(data.PS$WorkTimeMax_PS))

mod.work.time <- brm(WorkTimeMax_PS ~ OffCamera + (1 | Species),
                   family = lognormal(),  data = data.PS, # lognormal
                   chains = 4,  cores = 4,  iter = 4000,
                   prior = base_priors_ln,
                   control = list(adapt_delta = 0.99, max_treedepth=15))

# Activity rate
hist(log(data.PS$ActivityRateMax_PS))

mod.act.rate <- brm(ActivityRateMax_PS ~ OffCamera + (1 | Species),
                   family = lognormal(),  data = data.PS, # lognormal
                   chains = 4,  cores = 4,  iter = 4000,
                   prior = base_priors_ln,
                   control = list(adapt_delta = 0.99, max_treedepth=15))

# Solve success
mod.solve <- brm(Solve_PS_Binary ~ (1 | Species),
                    family = bernoulli(link = "logit"), data = data.PS, # bernoulli
                    chains = 4,  cores = 4,  iter = 4000,
                    prior = c(prior(normal(0, 1.5), class = "Intercept"),
                              prior(exponential(1), class = "sd")),
                   control = list(adapt_delta = 0.99, max_treedepth=15))


# Save models
saveRDS(mod.prop.vig, "Model_Output/Q1-First-Attempts/mod.prop.vig.rds")
saveRDS(mod.look.cam, "Model_Output/Q1-First-Attempts/mod.look.cam.rds")
saveRDS(mod.touch, "Model_Output/Q1-First-Attempts/mod.touch.rds")
saveRDS(mod.tot.exp, "Model_Output/Q1-First-Attempts/mod.tot.exp.rds")
saveRDS(mod.exp.div, "Model_Output/Q1-First-Attempts/mod.exp.div.rds")
saveRDS(mod.work.time, "Model_Output/Q1-First-Attempts/mod.work.time.rds")
saveRDS(mod.act.rate, "Model_Output/Q1-First-Attempts/mod.act.rate.rds")
saveRDS(mod.solve, "Model_Output/Q1-First-Attempts/mod.solve.rds")
```

### Load Models

If not re-running model, load all model output:

```{r}
# Load models
model_files <- list.files("Model_Output/Q1-First-Attempts/", pattern = "\\.rds$", full.names = TRUE)

# Load each model to the environment
walk(model_files, function(file) {
  model_name <- str_remove(basename(file), "\\.rds$")
  assign(model_name, readRDS(file), envir = .GlobalEnv)
})
```

### Assess model fit:

```{r}
# Posterior predictive checks
pp_check(mod.prop.vig, ndraws=100)
pp_check(mod.look.cam, ndraws=100)
pp_check(mod.touch, ndraws=100)
pp_check(mod.tot.exp, ndraws=100)
pp_check(mod.exp.div, ndraws=100)
pp_check(mod.work.time, ndraws=100)
pp_check(mod.act.rate, ndraws=100)
pp_check(mod.solve, ndraws=100)


# DHARMA
dh_check_brms(mod.prop.vig)
dh_check_brms(mod.look.cam)
dh_check_brms(mod.touch)
dh_check_brms(mod.tot.exp)
dh_check_brms(mod.exp.div)
dh_check_brms(mod.work.time)
dh_check_brms(mod.act.rate)
dh_check_brms(mod.solve)
```

# Extract and Format Model Results

```{r}
# Extract tables for each model
table.prop.vig <- broom.mixed::tidy(mod.prop.vig, conf.method="HPDinterval") %>% 
  mutate(term = c('Conditional Intercept', 'Zero-Inflation Intercept', 'Off Camera: Yes', 'Conditional Species Intercept', 'Zero-Inflation Species Intercept')) %>% 
  rename_all(str_to_title) %>% 
  select(-c(Component, Group, Std.error)) %>% 
  mutate(Model = 'Proportion Vigilant') %>% 
  relocate(Model)

table.look.cam <- broom.mixed::tidy(mod.look.cam, conf.method="HPDinterval") %>% 
  mutate(term = c('Intercept', 'Species Intercept')) %>% 
  rename_all(str_to_title) %>% 
  select(-c(Component, Group, Std.error)) %>% 
  mutate(Model = 'Look At Camera') %>% 
  relocate(Model)

table.touch <- broom.mixed::tidy(mod.touch, conf.method="HPDinterval") %>% 
  mutate(term = c('Intercept',  'Species Intercept')) %>% 
  rename_all(str_to_title) %>% 
  select(-c(Component, Group, Std.error)) %>% 
  mutate(Model = 'Touch Camera') %>% 
  relocate(Model)

table.tot.exp <- broom.mixed::tidy(mod.tot.exp, conf.method="HPDinterval") %>% 
  mutate(term = c('Intercept', 'Off Camera: Yes', 'Species Intercept')) %>%
  rename_all(str_to_title) %>% 
  select(-c(Component, Group, Std.error)) %>% 
  mutate(Model = 'Total Exploration') %>% 
  relocate(Model)

table.exp.div <- broom.mixed::tidy(mod.exp.div, conf.method="HPDinterval") %>% 
  mutate(term = c('Intercept', 'Off Camera: Yes', 'Species Intercept')) %>% 
  rename_all(str_to_title) %>% 
  select(-c(Component, Group, Std.error)) %>% 
  mutate(Model = 'Exploratory Diversity') %>% 
  relocate(Model)

table.work.time <- broom.mixed::tidy(mod.work.time, conf.method="HPDinterval") %>% 
  mutate(term = c('Intercept', 'Off Camera: Yes', 'Species Intercept', 'Residual')) %>% 
  rename_all(str_to_title) %>% 
  select(-c(Component, Group, Std.error)) %>% 
  mutate(Model = 'Work Time') %>% 
  relocate(Model)

table.act.time <- broom.mixed::tidy(mod.act.rate, conf.method="HPDinterval") %>% 
  mutate(term = c('Intercept', 'Off Camera: Yes', 'Species Intercept', 'Residual')) %>% 
  rename_all(str_to_title) %>% 
  select(-c(Component, Group, Std.error)) %>% 
  mutate(Model = 'Activity Rate') %>% 
  relocate(Model)

table.solve <- broom.mixed::tidy(mod.solve, conf.method="HPDinterval") %>% 
  mutate(term = c('Intercept', 'Species Intercept')) %>% 
  rename_all(str_to_title) %>% 
  select(-c(Component, Group, Std.error)) %>% 
  mutate(Model = 'Solve Success') %>% 
  relocate(Model)

# Calculate R2 values
r2_propvig <- performance::r2_bayes(mod.prop.vig)
r2_lookcam <- performance::r2_bayes(mod.look.cam)
r2_touch <- performance::r2_bayes(mod.touch)
r2_totexp <- performance::r2_bayes(mod.tot.exp)
r2_expdiv <- performance::r2_bayes(mod.exp.div)
r2_worktime <- performance::r2_bayes(mod.work.time)
r2_actrate <- performance::r2_bayes(mod.act.rate)
r2_solve <- performance::r2_bayes(mod.solve)

# Store in a list
r2_list <- list(
  'Proportion Vigilant' = r2_propvig,
  'Look At Camera' = r2_lookcam,
  'Touch Camera' = r2_touch,
  'Total Exploration' = r2_totexp,
  'Exploratory Diversity' = r2_expdiv, 
  'Work Time' = r2_worktime,
  'Activity Rate' = r2_actrate,
  'Solve Success' = r2_solve)

r2_table <- data.frame(
  Model = names(r2_list),
  Marginal_R2 = sapply(r2_list, function(x) x$R2_Bayes_marginal),
  Conditional_R2 = sapply(r2_list, function(x) x$R2_Bayes),
  Marginal_CI_low = sapply(r2_list, function(x) attr(x, "CI")$R2_Bayes_marginal$CI_low),
  Marginal_CI_high = sapply(r2_list, function(x) attr(x, "CI")$R2_Bayes_marginal$CI_high),
  Conditional_CI_low = sapply(r2_list, function(x) attr(x, "CI")$R2_Bayes$CI_low),
  Conditional_CI_high = sapply(r2_list, function(x) attr(x, "CI")$R2_Bayes$CI_high)) %>% 
  mutate('Marginal R2 [95% CI]' = paste0(
    sprintf("%.2f", Marginal_R2), " [",
    sprintf("%.2f", Marginal_CI_low), ", ",
    sprintf("%.2f", Marginal_CI_high), "]"
  )) %>% 
  mutate('Conditional R2 [95% CI]' = paste0(
    sprintf("%.2f", Conditional_R2), " [",
    sprintf("%.2f", Conditional_CI_low), ", ",
    sprintf("%.2f", Conditional_CI_high), "]"
  )) %>% 
  select(c(Model, 'Marginal R2 [95% CI]', 'Conditional R2 [95% CI]'))

# bind tables
big_table <- bind_rows(table.prop.vig, table.look.cam, table.touch, table.tot.exp, table.exp.div, table.work.time, table.act.time, table.solve) %>% 
  
  # Rename fixed and random effects
  mutate(Effect = recode(Effect,
                         "fixed" = "Fixed",
                         "ran_pars" = "Random")) %>% 
  
  # Add asterisk if CI does NOT overlap zero
  mutate(Sig = ifelse(Conf.low > 0 | Conf.high < 0, "*", "")) %>%
  
  # Combine estimate and 95% HPD
  mutate('Estimate [95% HPD]' = paste0(
    sprintf("%.2f", Estimate), " [",
    sprintf("%.2f", Conf.low), ", ",
    sprintf("%.2f", Conf.high), "]"
  )) %>%
  select(-Estimate, -Conf.low, -Conf.high) %>% 
  
  # Add R2 values
  left_join(r2_table, by="Model")

# Save model output
# write.csv(big_table, "Model_Output/Q1-First-Attempts/Q1_results_table.csv")
```

# Plot random intercepts

### Proportion Vigilant

```{r}
# Extract posterior samples
post_samps <- posterior_samples(mod.prop.vig)

# Extract species random intercepts for mean part (mu)
mu_species_cols <- grep("^r_Species\\[.*Intercept\\]$", colnames(post_samps), value = TRUE)

posterior_species_mu <- post_samps[, mu_species_cols] %>%
  as_tibble() %>%
  pivot_longer(cols = everything(), names_to = "species_id", values_to = "value") %>%
  mutate(Species = sub("r_Species\\[(.*),Intercept\\]", "\\1", species_id))

# Add draw number to track posterior draws
posterior_species_mu <- posterior_species_mu %>%
  mutate(draw = rep(1:(nrow(post_samps)), each = length(mu_species_cols)))

# Add fixed intercept for mu (population-level intercept)
fixef_draws_mu <- as.data.frame(post_samps$b_Intercept) %>% 
  mutate(draw = seq(1:nrow(post_samps)))

posterior_species_mu <- posterior_species_mu %>%
  left_join(fixef_draws_mu, by="draw") %>% 
  rename(b_Intercept = 'post_samps$b_Intercept') %>% 
  mutate(value_resp = b_Intercept + value) %>%
  mutate(mu = plogis(value_resp))   # inverse logit link for mean

# Extract species random intercepts for zero-inflation (zi) part
zi_species_cols <- grep("^r_Species__zi\\[.*Intercept\\]$", colnames(post_samps), value = TRUE)

posterior_species_zi <- post_samps[, zi_species_cols] %>%
  as_tibble() %>%
  pivot_longer(cols = everything(), names_to = "species_id", values_to = "value") %>%
  mutate(Species = sub("r_Species__zi\\[(.*),Intercept\\]", "\\1", species_id))

posterior_species_zi <- posterior_species_zi %>%
  mutate(draw = rep(1:(nrow(post_samps)), each = length(zi_species_cols)))

# Add fixed intercept for zero inflation (population-level zi intercept)
fixef_draws_zi <- as.data.frame(post_samps$b_zi_Intercept) %>% 
  mutate(draw = seq(1:nrow(post_samps)))

posterior_species_zi <- posterior_species_zi %>%
  left_join(fixef_draws_zi, by="draw") %>% 
  rename(b_zi_Intercept = 'post_samps$b_zi_Intercept') %>% 
  mutate(value_zi = b_zi_Intercept + value) %>%
  mutate(zi_prob = plogis(value_zi))

# Combine mu and zi by Species and draw to get probability of vigilance while incorporating zero inflation
posterior_combined <- posterior_species_mu %>%
  inner_join(posterior_species_zi, by = c("Species", "draw")) %>%
  mutate(expected_prop = (1 - zi_prob) * mu) # probability of being vigilant times mean proportion vigilant


# Calculate mean expected proportion for each species and CI
posterior_plotting <- posterior_combined %>%
  group_by(Species) %>%
  summarise(
    mean_prob = mean(expected_prop),
    lower_95 = HPDinterval(as.mcmc(expected_prop), prob = 0.95)[,1],
    upper_95 = HPDinterval(as.mcmc(expected_prop), prob = 0.95)[,2],
    lower_90 = HPDinterval(as.mcmc(expected_prop), prob = 0.90)[,1],
    upper_90 = HPDinterval(as.mcmc(expected_prop), prob = 0.90)[,2]) %>% 
  ungroup()

p.vigilance <- ggplot(posterior_plotting, aes(y = Species, x = mean_prob, color = Species, fill=Species)) +
  # 95% interval (transparent)
  geom_linerange(aes(xmin = lower_95, xmax = upper_95), alpha = 0.6, size = 1) +
  # 90% interval (solid)
  geom_linerange(aes(xmin = lower_90, xmax = upper_90), alpha = 1, size = 1.5) +
  # Mean point
  geom_point(shape = 124, size = 4, color = "black") +
  scale_x_continuous(limits = c(0, 1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_y_discrete(labels = c("Coyote", "Domestic Cat", "Opossum", "Raccoon - Alone",
                              "Raccoon - Group", "Skunk"))+
  labs(x = "Predicted Proportion Time Vigilant",
       y = NULL) +
  scale_color_manual(values = colors) +
  theme_custom() +
  theme(legend.position = "none")

p.vigilance
```

### Look at Camera

```{r}
# Extract posterior samples
post_samps <- posterior_samples(mod.look.cam)

# Extract species random intercepts
species_cols <- grep("^r_Species\\[.*Intercept\\]$", colnames(post_samps), value = TRUE)

posterior_species <- post_samps[, species_cols] %>%
  as_tibble() %>%
  pivot_longer(cols = everything(), names_to = "species_id", values_to = "value") %>%
  mutate(Species = sub("r_Species\\[(.*),Intercept\\]", "\\1", species_id))

# Add draw number to track posterior draws
posterior_species <- posterior_species %>%
  mutate(draw = rep(1:(nrow(post_samps)), each = length(species_cols)))

# Add fixed intercept (population-level intercept)
fixef_draws <- as.data.frame(post_samps$b_Intercept) %>% 
  mutate(draw = seq(1:nrow(post_samps)))

posterior_species <- posterior_species %>%
  left_join(fixef_draws, by="draw") %>% 
  rename(b_Intercept = 'post_samps$b_Intercept') %>% 
  mutate(value_resp =  b_Intercept + value) %>%
  mutate(prob = plogis(value_resp)) # inverse logit to get probability

# Calculate mean expected proportion for each species and CI
posterior_plotting <- posterior_species %>%
  group_by(Species) %>%
  summarise(
    mean_prob = mean(prob),
    lower_95 = HPDinterval(as.mcmc(prob), prob = 0.95)[,1],
    upper_95 = HPDinterval(as.mcmc(prob), prob = 0.95)[,2],
    lower_90 = HPDinterval(as.mcmc(prob), prob = 0.90)[,1],
    upper_90 = HPDinterval(as.mcmc(prob), prob = 0.90)[,2]) %>% 
  ungroup()


# Plot
p.look.cam <- ggplot(posterior_plotting, aes(y = Species, x = mean_prob, color = Species, fill=Species)) +
  # 95% interval (transparent)
  geom_linerange(aes(xmin = lower_95, xmax = upper_95), alpha = 0.6, size = 1) +
  # 90% interval (solid)
  geom_linerange(aes(xmin = lower_90, xmax = upper_90), alpha = 1, size = 1.5) +
  # Mean point
  geom_point(shape = 124, size = 4, color = "black") +
  scale_x_continuous(limits = c(0, 1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_y_discrete(labels = c("Coyote", "Domestic Cat", "Opossum", "Raccoon - Alone",
                              "Raccoon - Group", "Skunk"))+
  labs(x = "Predicted Probability of Looking at Camera",
       y = NULL) +
  scale_color_manual(values = colors) +
  theme_custom() +
  theme(legend.position = "none")

p.look.cam
```

### Touch puzzle

```{r}
# Extract posterior samples for mod.touch
post_samps <- posterior_samples(mod.touch)

# Extract species random intercepts columns
posterior_species <- post_samps[, species_cols] %>%
  as_tibble() %>%
  pivot_longer(cols = everything(), names_to = "species_id", values_to = "value") %>%
  mutate(Species = sub("r_Species\\[(.*),Intercept\\]", "\\1", species_id))

# Add draw number to track posterior draws
posterior_species <- posterior_species %>%
  mutate(draw = rep(1:(nrow(post_samps)), each = length(species_cols)))

# Add fixed intercept (population-level intercept)
fixef_draws <- as.data.frame(post_samps$b_Intercept) %>% 
  mutate(draw = seq(1:nrow(post_samps)))

posterior_species <- posterior_species %>%
  left_join(fixef_draws, by="draw") %>% 
  rename(b_Intercept = 'post_samps$b_Intercept') %>% 
  mutate(value_resp = b_Intercept + value) %>%
  mutate(prob = plogis(value_resp)) # inverse logit to get probability

# Calculate mean expected proportion for each species and CI
posterior_plotting <- posterior_species %>%
  group_by(Species) %>%
  summarise(
    mean_prob = mean(prob),
    lower_95 = HPDinterval(as.mcmc(prob), prob = 0.95)[,1],
    upper_95 = HPDinterval(as.mcmc(prob), prob = 0.95)[,2],
    lower_90 = HPDinterval(as.mcmc(prob), prob = 0.90)[,1],
    upper_90 = HPDinterval(as.mcmc(prob), prob = 0.90)[,2]) %>% 
  ungroup()

# Plot
p.touch.puz <- ggplot(posterior_plotting, aes(y = Species, x = mean_prob, color = Species, fill=Species)) +
  # 95% interval (transparent)
  geom_linerange(aes(xmin = lower_95, xmax = upper_95), alpha = 0.6, size = 1) +
  # 90% interval (solid)
  geom_linerange(aes(xmin = lower_90, xmax = upper_90), alpha = 1, size = 1.5) +
  # Mean point
  geom_point(shape = 124, size = 4, color = "black") +
  scale_x_continuous(limits = c(0, 1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_y_discrete(labels = c("Coyote", "Domestic Cat", "Opossum", "Raccoon - Alone",
                              "Raccoon - Group", "Skunk"))+
  labs(x = "Predicted Probability of Touching Puzzle",
       y = NULL) +
  scale_color_manual(values = colors) +
  theme_custom() +
  theme(legend.position = "none")

p.touch.puz
```

### Total exploration

```{r}
# Extract posterior samples
post_samps <- posterior_samples(mod.tot.exp)

# Extract species random intercept columns for mean (mu)
posterior_species <- post_samps[, species_cols] %>%
  as_tibble() %>%
  pivot_longer(cols = everything(), names_to = "species_id", values_to = "value") %>%
  mutate(Species = sub("r_Species\\[(.*),Intercept\\]", "\\1", species_id))

# Add draw number to track posterior draws
posterior_species <- posterior_species %>%
  mutate(draw = rep(1:(nrow(post_samps)), each = length(species_cols)))

# Add fixed intercept (population-level intercept)
fixef_draws <- as.data.frame(post_samps$b_Intercept) %>% 
  mutate(draw = seq(1:nrow(post_samps)))

posterior_species <- posterior_species %>%
  left_join(fixef_draws, by="draw") %>% 
  rename(b_Intercept = 'post_samps$b_Intercept') %>% 
  mutate(value_resp = b_Intercept + value) %>%
  mutate(mu = exp(value_resp)) # inverse log link to get expected counts

# Calculate mean expected proportion for each species and CI
posterior_plotting <- posterior_species %>%
  group_by(Species) %>%
  summarise(
    mean_mu = mean(mu),
    lower_95 = HPDinterval(as.mcmc(mu), prob = 0.95)[,1],
    upper_95 = HPDinterval(as.mcmc(mu), prob = 0.95)[,2],
    lower_90 = HPDinterval(as.mcmc(mu), prob = 0.90)[,1],
    upper_90 = HPDinterval(as.mcmc(mu), prob = 0.90)[,2]) %>% 
  ungroup()


# Plot
p.tot.exp <- ggplot(posterior_plotting, aes(y = Species, x = mean_mu, color = Species, fill=Species)) +
  # 95% interval (transparent)
  geom_linerange(aes(xmin = lower_95, xmax = upper_95), alpha = 0.6, size = 1) +
  # 90% interval (solid)
  geom_linerange(aes(xmin = lower_90, xmax = upper_90), alpha = 1, size = 1.5) +
  # Mean point
  geom_point(shape = 124, size = 4, color = "black") +
  scale_x_continuous(limits = c(0, 22.5),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_y_discrete(labels = c("Coyote", "Domestic Cat", "Opossum", "Raccoon - Alone",
                              "Raccoon - Group", "Skunk"))+
  labs(x = "Predicted Total Exploration (Count)",
       y = NULL) +
  scale_color_manual(values = colors) +
  theme_custom() +
  theme(legend.position = "none")

p.tot.exp
```

### Exploratory diversity

```{r}
# Extract posterior samples
post_samps <- posterior_samples(mod.exp.div)
colnames(post_samps)

# Extract species random intercept columns for mean (mu)
species_cols <- grep("^r_Species\\[.*Intercept\\]$", colnames(post_samps), value = TRUE)

posterior_species <- post_samps[, species_cols] %>%
  as_tibble() %>%
  pivot_longer(cols = everything(), names_to = "species_id", values_to = "value") %>%
  mutate(Species = sub("r_Species\\[(.*),Intercept\\]", "\\1", species_id))

# Add draw number to track posterior draws
posterior_species <- posterior_species %>%
  mutate(draw = rep(1:(nrow(post_samps)), each = length(species_cols)))

# Add fixed intercept (population-level intercept)
fixef_draws <- as.data.frame(post_samps$b_Intercept) %>% 
  mutate(draw = seq(1:nrow(post_samps)))

posterior_species <- posterior_species %>%
  left_join(fixef_draws, by="draw") %>% 
  rename(b_Intercept = 'post_samps$b_Intercept') %>% 
  mutate(value_resp = b_Intercept +  value) %>%
  mutate(mu = exp(value_resp)) # exponentiate to get expected counts

# Calculate mean expected proportion for each species and CI
posterior_plotting <- posterior_species %>%
  group_by(Species) %>%
  summarise(
    mean_mu = mean(mu),
    lower_95 = HPDinterval(as.mcmc(mu), prob = 0.95)[,1],
    upper_95 = HPDinterval(as.mcmc(mu), prob = 0.95)[,2],
    lower_90 = HPDinterval(as.mcmc(mu), prob = 0.90)[,1],
    upper_90 = HPDinterval(as.mcmc(mu), prob = 0.90)[,2]) %>% 
  ungroup()

# Plot
p.exp.div <- ggplot(posterior_plotting, aes(y = Species, x = mean_mu, color = Species, fill=Species)) +
  # 95% interval (transparent)
  geom_linerange(aes(xmin = lower_95, xmax = upper_95), alpha = 0.6, size = 1) +
  # 90% interval (solid)
  geom_linerange(aes(xmin = lower_90, xmax = upper_90), alpha = 1, size = 1.5) +
  # Mean point
  geom_point(shape = 124, size = 4, color = "black") +
  scale_x_continuous(limits = c(0, 7),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_y_discrete(labels = c("Coyote", "Domestic Cat", "Opossum", "Raccoon - Alone",
                              "Raccoon - Group", "Skunk"))+
  labs(x = "Predicted Exploratory Diversity (Count)",
       y = NULL) +
  scale_color_manual(values = colors) +
  theme_custom() +
  theme(legend.position = "none")

p.exp.div
```

### Work time

```{r}
# Extract posterior samples
post_samps <- posterior_samples(mod.work.time)

# Extract species random intercept columns for mean (mu)
posterior_species <- post_samps[, species_cols] %>%
  as_tibble() %>%
  pivot_longer(cols = everything(), names_to = "species_id", values_to = "value") %>%
  mutate(Species = sub("r_Species\\[(.*),Intercept\\]", "\\1", species_id))

# Add draw number to track posterior draws
posterior_species <- posterior_species %>%
  mutate(draw = rep(1:(nrow(post_samps)), each = length(species_cols)))

# Add fixed intercept (population-level intercept)
fixef_draws <- as.data.frame(post_samps$b_Intercept) %>% 
  mutate(draw = seq(1:nrow(post_samps)))

posterior_species <- posterior_species %>%
  left_join(fixef_draws, by="draw") %>% 
  rename(b_Intercept = 'post_samps$b_Intercept') %>% 
  mutate(mu = exp(b_Intercept + value))

# Calculate mean expected proportion for each species and CI
posterior_plotting <- posterior_species %>%
  group_by(Species) %>%
  summarise(
    mean_mu = mean(mu),
    lower_95 = HPDinterval(as.mcmc(mu), prob = 0.95)[,1],
    upper_95 = HPDinterval(as.mcmc(mu), prob = 0.95)[,2],
    lower_90 = HPDinterval(as.mcmc(mu), prob = 0.90)[,1],
    upper_90 = HPDinterval(as.mcmc(mu), prob = 0.90)[,2]) %>% 
  ungroup()

# Plot
p.work.time <- ggplot(posterior_plotting, aes(y = Species, x = mean_mu, color = Species, fill=Species)) +
  # 95% interval (transparent)
  geom_linerange(aes(xmin = lower_95, xmax = upper_95), alpha = 0.6, size = 1) +
  # 90% interval (solid)
  geom_linerange(aes(xmin = lower_90, xmax = upper_90), alpha = 1, size = 1.5) +
  # Mean point
  geom_point(shape = 124, size = 4, color = "black") +
  scale_x_continuous(limits = c(0, 76),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_y_discrete(labels = c("Coyote", "Domestic Cat", "Opossum", "Raccoon - Alone",
                              "Raccoon - Group", "Skunk"))+
  labs(x = "Predicted Work Time (Seconds)",
       y = NULL) +
  scale_color_manual(values = colors) +
  theme_custom() +
  theme(legend.position = "none")

p.work.time
```

### Activity rate

```{r}
# Extract posterior samples
post_samps <- posterior_samples(mod.act.rate)

# Extract species random intercept columns for mean (mu)
posterior_species <- post_samps[, species_cols] %>%
  as_tibble() %>%
  pivot_longer(cols = everything(), names_to = "species_id", values_to = "value") %>%
  mutate(Species = sub("r_Species\\[(.*),Intercept\\]", "\\1", species_id))

# Add draw number to track posterior draws
posterior_species <- posterior_species %>%
  mutate(draw = rep(1:(nrow(post_samps)), each = length(species_cols)))

# Add fixed intercept (population-level intercept)
fixef_draws <- as.data.frame(post_samps$b_Intercept) %>% 
  mutate(draw = seq(1:nrow(post_samps)))

posterior_species <- posterior_species %>%
  left_join(fixef_draws, by="draw") %>% 
  rename(b_Intercept = 'post_samps$b_Intercept') %>% 
  mutate(mu = exp(b_Intercept + value))

# Calculate mean expected proportion for each species and CI
posterior_plotting <- posterior_species %>%
  group_by(Species) %>%
  summarise(
    mean_mu = mean(mu),
    lower_95 = HPDinterval(as.mcmc(mu), prob = 0.95)[,1],
    upper_95 = HPDinterval(as.mcmc(mu), prob = 0.95)[,2],
    lower_90 = HPDinterval(as.mcmc(mu), prob = 0.90)[,1],
    upper_90 = HPDinterval(as.mcmc(mu), prob = 0.90)[,2]) %>% 
  ungroup()

# 6. Plot ridge plot with mean expected counts points
p.act.time <- ggplot(posterior_plotting, aes(y = Species, x = mean_mu, color = Species, fill=Species)) +
  # 95% interval (transparent)
  geom_linerange(aes(xmin = lower_95, xmax = upper_95), alpha = 0.6, size = 1) +
  # 90% interval (solid)
  geom_linerange(aes(xmin = lower_90, xmax = upper_90), alpha = 1, size = 1.5) +
  # Mean point
  geom_point(shape = 124, size = 4, color = "black") +
  scale_x_continuous(limits = c(0, 1.5),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_y_discrete(labels = c("Coyote", "Domestic Cat", "Opossum", "Raccoon - Alone",
                              "Raccoon - Group", "Skunk"))+
  labs(x = "Predicted Activity Rate",
       y = NULL) +
  scale_color_manual(values = colors) +
  theme_custom() +
  theme(legend.position = "none")

p.act.time
```

### Problem-solving success

```{r}
# Extract posterior samples
post_samps <- posterior_samples(mod.solve)

# Extract species random intercepts columns
posterior_species <- post_samps[, species_cols] %>%
  as_tibble() %>%
  pivot_longer(cols = everything(), names_to = "species_id", values_to = "value") %>%
  mutate(Species = sub("r_Species\\[(.*),Intercept\\]", "\\1", species_id))

# Add draw number to track posterior draws
posterior_species <- posterior_species %>%
  mutate(draw = rep(1:(nrow(post_samps)), each = length(species_cols)))

# Add fixed intercept (population-level intercept)
fixef_draws <- as.data.frame(post_samps$b_Intercept) %>% 
  mutate(draw = seq(1:nrow(post_samps)))

posterior_species <- posterior_species %>%
  left_join(fixef_draws, by="draw") %>% 
  rename(b_Intercept = 'post_samps$b_Intercept') %>% 
  mutate(value_resp = b_Intercept + value) %>%
  mutate(prob = plogis(value_resp)) # inverse logit to get probability

# Calculate mean expected proportion for each species and CI
posterior_plotting <- posterior_species %>%
  group_by(Species) %>%
  summarise(
    mean_prob = mean(prob),
    lower_95 = HPDinterval(as.mcmc(prob), prob = 0.95)[,1],
    upper_95 = HPDinterval(as.mcmc(prob), prob = 0.95)[,2],
    lower_90 = HPDinterval(as.mcmc(prob), prob = 0.90)[,1],
    upper_90 = HPDinterval(as.mcmc(prob), prob = 0.90)[,2]) %>% 
  ungroup()

# 6. Plot ridge plot with mean probability points
p.solve <- ggplot(posterior_plotting, aes(y = Species, x = mean_prob, color = Species, fill=Species)) +
  # 95% interval (transparent)
  geom_linerange(aes(xmin = lower_95, xmax = upper_95), alpha = 0.6, size = 1) +
  # 90% interval (solid)
  geom_linerange(aes(xmin = lower_90, xmax = upper_90), alpha = 1, size = 1.5) +
  # Mean point
  geom_point(shape = 124, size = 4, color = "black") +
  scale_x_continuous(limits = c(0, 1),
                     expand = expansion(mult = c(0.02, 0.02))) +
  scale_y_discrete(labels = c("Coyote", "Domestic Cat", "Opossum", "Raccoon - Alone",
                              "Raccoon - Group", "Skunk"))+
  labs(x = "Predicted Probability of Solving",
       y = NULL) +
  scale_color_manual(values = colors) +
  theme_custom() +
  theme(legend.position = "none")

p.solve
```

## Arrange plots

```{r}
# Arrange risk-taking plots
ggarrange(p.vigilance, p.look.cam, p.touch.puz, p.tot.exp,
          align="hv")

# ggsave("Figures/RT.plots_first_attempts.png", dpi=600, height=6, width=10)

# Arrange problem-solving plots
ggarrange(p.exp.div, p.work.time, p.act.time, p.solve,
          align="hv")

# ggsave("Figures/PS.plots_first_attempts.png", dpi=600, height=6, width=10)
```
